---
title: "Stat 184 Final Project"
author: "Jessica Dougherty"
date: "Due Dec. 13th"
output: html_notebook
---

Guiding Question: What social media platforms do each political party use most frequently?
Sub-question: Which social media platform does each party have a bigger following on? 

This data source is from the Pew Research Center. Pew Research Center collected all of the Facebook posts and tweets posted by all official and unofficial account maintained by every voting member of the U.S. Senate and House of Representatives between Jan. 1, 2015, and May 31, 2020. This data was originally collected to study political rhetoric and attempt to build a continually updated database of political social media activity across platforms such as Facebook and Twitter. One case is one member of congress's social media stats. This includes party affiliation, number of followers, posts, average reactions, average shares, etc. There are 1078 cases in the data set. I plan to use the variables platform, party, total posts, average favorites, likes, retweets, and shares. I would like to note that in the data Independents are associated with the party they caucused with for the majority of the study period, but identified with parentheses. I will be considering them as the party they caucused with for the purpose of this assignment.

For my second data set, I am using a set from the same study by Pew Research Center. This data set looks at the percentage change of social media activity when big events occur. It looks at the top ten events for both parties. Some of the variables include date, party affiliation, percentage change of likes/shares on Facebook/Twitter, etc. It is kept and updated via the same pollster and study.

Load packages
```{r}
library(ggplot2)
library(tidyverse)
```

Import Data Sets
```{r}
library(readxl)
Social_media_statistics_for_members_of_the_116th_Congress <- read_excel("Social media statistics for members of the 116th Congress.xlsx", 
    skip = 1)
View(Social_media_statistics_for_members_of_the_116th_Congress)

library(readr)
Top_10_events_associated_with_increased_engagement_in_each_party <- read_csv("appendix csvs/Top 10 events associated with increased engagement in each party.csv")
View(Top_10_events_associated_with_increased_engagement_in_each_party)
```

Summary Stats
```{r}
smStats <- Social_media_statistics_for_members_of_the_116th_Congress
View(smStats)

smStats%>%
  summarise(n = n(), meanLikes = mean(`Average Post Favorites/Reactions`), meanShares = mean(`Average Post Retweets/Shares`))
```

Over all 1078 members of congress in this study, the average number of likes or reactions they recieve is 641. The average number of shares in 162. For reference, in 2013 Kim Kardashian recieved about 600 retweets within 50 minutes of posting something. This is not a lot of likes or shares for a "famous person". 

Boxplot of Favorites/Reactions
```{r}
ggplot(data = smStats, aes(x = adjParty, y = `Average Post Favorites/Reactions`)) + geom_boxplot()+scale_y_continuous(limits = c(0, 1000))
```

Looking at this box plot of favorites and reactions among congress, the two parties are pretty similar numberwise. However, there is a slightly bigger range for the republican party. It is also interesting to note that there are a few democrats who's average favorites/reacts are higher than the highest republican.

Boxplot of Retweets/Shares
```{r}
ggplot(data = smStats, aes(x=adjParty, y=`Average Post Retweets/Shares`)) + geom_boxplot() + scale_y_continuous(limits = c(0, 300))
```

In terms of retweets and shares, the two parties seem to look pretty even. Again, the republican's have a bigger range, but not by much. 

Second Data Set Summary Stats
```{r}
topTen <- Top_10_events_associated_with_increased_engagement_in_each_party
View(topTen)

topTen%>%
  group_by(Party)%>%
  summarise(n=n(), meanChangeFBLikes = mean(`Percentage Change in Facebook Reactions`),
            meanChangeFBShares = mean(`Percentage Change in Facebook Shares`),
            meanChangeTwitRt = mean(`Percentage Change in Twitter Retweets`),
            meanChangeTwitFav = mean(`Percentage Change in Twitter Favorites`))
```
For this data set, we see the mean percent change in likes, shares, retweets, and favorites only fluxuate by less than 4%. However, when you look at the raw data set, the democrat's highest percetage is 12.617 (Twitter Favorites) while the republican's is 9.625 (FB Shares). 

Democrat Stats
```{r}
Dem <- smStats%>%
  filter(adjParty == "D")
```